---
title: "R6 basics for mlr3"
output: html_notebook
---

## Why mlr3?

It uses R6, that addresses shortcomings of earlier OO implementations e.g. S3 used in `mlr`.

## Quick R6 examples for beginners

If you have done any object-oriented programming before, R6 should feel familiar.

```{r}
library(R6)

Person <- R6Class("Person",
  public = list(
    name = NULL,
    hair = NULL,
    initialize = function(name = NA, hair = NA) {
      self$name <-name
      self$hair <- hair
      self$greet()
    },
    set_hair = function(val){
      self$hair <- val
    },
    greet = function(){
      cat(paste0("Hi, my name is ", self$name, ".\n"))
    }
  )
)
```

We instantiate the object with `$new()`. For example:

```{r}
hud <- Person$new("Hud", "black")
hud
```

We can access the fields as such:

```{r}
hud$hair
```

And set the argument:

```{r}
hud$hair = "blonde"
hud$hair
```

## Quick `data.table` examples

Many `mlr3` functions return `data.table`s which can conveniently be subset or combined with other outputs.

The interface is similar to `data.frame()`:

```{r}
library(data.table)
dt = data.table(x = 1:6, y = rep(letters[1:3], each = 2))
dt
```

`data.table`s can be used much like `data.frame`s, but they do provide additional functionality that makes complex operations easier. For example, data can be summarized by groups with the `[` operator:

```{r}
dt[, mean(x), by = "y"]
```
